A Convolutional Neural Network for Outsole Recognition\tnoteref{t1}

### 2019-06-25 paper/paper.Rnw
language edits in conclusion

% Discussion of weaknesses
Although model predictions are generally accurate, there are a number of systematic misclassifications. The model has substantial difficulty predicting images with low contrast because image features are obscured. Exploring methods of color correction, such as histogram equalization, may prove useful for eliminating the effect of contrast on predictions.

Another source of systematic misclassifications stems from the inherent overlap between features of common geometric shapes; since CNNs make predictions via bottom-up feature integration, it is more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles. This issue is exacerbated by high variability of those shapes on outsoles, which requires the model to form a flexible understanding of each shape category, making distinction between similar shapes even more difficult. These issues of overlap and variability are at least in part due to our decision to use the features used by practitioners directly; some of this may also result from our operational definitions of various shapes, which did not take into account the model's capabilities and initial training. While a \fix{full} remedy may not be possible, there may be some modifications to operational class definitions that we can make to optimize labels for model prediction.

\fix{Although the model is not perfect in its current form, as a first pass tool it is performing well and improvements seem promising. That most of the misclassifications are systematic shows that the model is able to learn features and apply rules consistently. There are possible modifications to the labeling scheme that may make model predictions align more closely to human-assigned labels, but model consistency is a much higher good than ``accuracy" (in the sense of agreeing with human-assigned labels). Examiners themselves may vary in how to label ambiguous features (e.g., the half-rounded quadrilaterals in \fix{the dc heatmap}), so a consistent classification method provides a more unified definition of certain geometric shapes and facilitates comparison of outsoles with similar geometric features. All in all, this model shows that CNNS are a useful tool for classifying a large number of outsoles within a common classification scheme and can ultimately help streamline the current process of sorting and comparing shoes by class characteristics.}

\mt{I think the conclusion currently says everything I want it to (cutting off a conv block doesn't seem to help make the point). Definitely can use some language edits.}

>>>>

% Discussion of weaknesses
Although model predictions are generally accurate, there are a number of systematic misclassifications. The model has substantial difficulty predicting images with low contrast because image features are obscured. Exploring methods of color correction, such as histogram equalization, may prove useful for eliminating the effect of contrast on predictions. \svp{Initial attempts at implementing histogram equalization suggest that the overall accuracy improvement is modest; the model fit to images which have been equalized had approximately 92.5\% validation accuracy, an improvement of 1.5\% over the default model described in this paper.}

Another source of systematic misclassifications stems from the inherent overlap between features of common geometric shapes\svp{. }CNNs make predictions via bottom-up feature integration; \svp{as a result,} it is more difficult for the model to distinguish between geometrically similar shapes, such as chevrons and triangles. This issue is exacerbated by high variability of some shapes \svp{commonly found in outsole patterns}, which requires the model to form a flexible understanding of each shape category \svp{and makes} distinction between similar shapes even more difficult. These issues of overlap and variability are at least in part due to our decision to use the features used by practitioners directly; some of this may also result from our operational definitions of various shapes, which \svp{were created before we had a complete understanding of the practical limitations of CNNs.\sout{did not take into account the model's capabilities and initial training.}} While a \fix{full} remedy may not be possible, there may be some modifications to operational class definitions that we can make to optimize labels for model prediction. \svp{In addition, we will continue to improve the labeling of shapes in our database with the goal of expanding the breadth and quantity of labeled images available for model training.}

\svp{Although there are many areas for potential improvement in future iterations of this model, we have demonstrated the effectiveness of transfer learning to provide an initial label for shoe outsole patterns. As} most of the misclassifications are systematic, it is clear that the model is able to learn features and apply rules consistently. \svp{We expect that future modifications to the labeling criteria and data processing pipeline will improve the model's accuracy, but the 91\% validation accuracy} \fix{really should be looking at test set accuracy, but I don't know that number off the top of my head...} \svp{is promising considering that we do not have estimates of the consistency of examiners' estimates on ambiguous features.}  \svp{Taken together, these factors } show that \svp{Convolutional Neural Networks} are a useful tool for classifying a large number of outsoles within a common classification scheme and can ultimately help streamline the current process of sorting and comparing shoes by class characteristics.

% \fix{Although the model is not perfect in its current form, as a first pass tool it is performing well and improvements seem promising. That most of the misclassifications are systematic shows that the model is able to learn features and apply rules consistently. There are possible modifications to the labeling scheme that may make model predictions align more closely to human-assigned labels, but model consistency is a much higher good than ``accuracy" (in the sense of agreeing with human-assigned labels). Examiners themselves may vary in how to label ambiguous features (e.g., the half-rounded quadrilaterals in \fix{the dc heatmap}), so a consistent classification method provides a more unified definition of certain geometric shapes and facilitates comparison of outsoles with similar geometric features. All in all, this model shows that CNNS are a useful tool for classifying a large number of outsoles within a common classification scheme and can ultimately help streamline the current process of sorting and comparing shoes by class characteristics.}
%
% \mt{I think the conclusion currently says everything I want it to (cutting off a conv block doesn't seem to help make the point). Definitely can use some language edits.}
